---
title: "SSCECON207 workshop I:<br>Introduction to R and ggplot"
date: "Spring 2022"
output: 
  github_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

# Learning outcomes

In this tutorial you will learn the steps needed to create basic figures in R.
You can use these skills to create a poster where you analyze empirical evidence on a policy intervention in the context of the IS-LM model.

You will learn how to import data into R, how to filter and select the data that you need, and how to make basic figures using the `ggplot2` package.

# First steps

This tutorial assumes that you have already installed R and RStudio, and have read [A (very) short introduction to R](https://github.com/ClaudiaBrauer/A-very-short-introduction-to-R/blob/master/documents/A\%20(very)\%20short\%20introduction\%20to\%20R.pdf) and [How to make any plot in ggplot2?](http://r-statistics.co/ggplot2-Tutorial-With-R.html#6.1\%20Make\%20a\%20time\%20series\%20plot\%20(using\%20ggfortify)).

## Your working directory

Once you open RStudio, it is useful to set your working directory.
This is the folder where you will save your datafiles, R scripts and figures.
To set a folder as your working directors, copy the following code and change the file path to the folder that you would like to use.
If you are using Windows, you need to manually change the backslashes (\\) to forward slashes (/) before you run the code.

You can set the working directory with the following function:

```{r, eval = FALSE}
setwd("C:/Users/Tina/Documents/SSCECON207")
```

## Installing `tidyverse`

Next, you need to install and load the `tidyverse` package.
This is a collection of packages that allow clean workflows in R.
The `ggplot2` package that you will use to create figures is one of the included packages.

You need to install the package only once, but you need to load it every time you open and use R.
It is good practice to load the package on the top of each script.

```{r, eval = FALSE}
install.packages("tidyverse") # install the package
```
```{r}
library(tidyverse) # load the package
```

# Importing and cleaning the data

## Importing a CSV file

Next, you need to import your data, which is in the format of a CSV (comma-separated values) file.
Download the data file from Moodle (add filename, location) and save it in the same folder that you set as your working directory.

You can import a file using the `read_csv()` function and by assigning the file to an object.

```{r}
data <- read_csv("ISLM_data.csv")
```

## Filtering the data

For your project you'll only need to use part of the data.
To see what the data frame looks like, you can view it by clicking on the name of the object in the Environment tab in the top right corner of RStudio.

The file "available_data.csv" provides a list of the countries and time periods available in the dataset.
You can use this file to select a policy intervention to study.
You can view the file [here](https://github.com/ucrdatacenter/projects/blob/main/SSCECON207/available_data.csv).

Once you decide which country and time period you would like to use in your project, you need to filter the data to remove the observations that you don't need.
You can do that with the `filter()` function.
For example if you are looking for US data between 2006-2013, you can do that as follows:

```{r}
data_US <- data %>%
  filter(year >= 2006, year <= 2013, code == "USA")
```

The pipe operator (` %>% `) means that the next function uses the previous result as an input: in this case, it takes the object `data` as the first argument of the `filter()` function.

The double equal sign in `code == "USA"` means that instead of assigning the string `"USA"` to a variable named country (which is what the code `code == "USA"` would do), you are testing whether the contents of the object `code` and the string `"USA"` are equal to each other.
Similarly, the less-than-or-equal and more-than-or-equal signs test the relationship between the contents of the variable `year` and the specified year.
If the result of the test is true, the function keeps the observation, and if the result is false, the observation is removed.
You can filter for country using both the variable `country`, which has full country names, or using the variable `code` with country code.
The results of the two operations will be equivalent.
Make sure to remember that R is case-sensitive, so e.g. a filter set as `country == "usa"` will not work.

## Other data manipulation

select & pivot_longer

# First `ggplot` figures

Figures made with `ggplot` are built from several layers.
You always use the same basic code structure to create a wide range of figures:

1. The `ggplot()` function creates a blank canvas for you to work on.
1. Geoms add the visual elements, such as points, lines, bars or other shapes.
1. Other specifications can include changing axis settings, setting the theme, adding labels, etc.
1. You connect all these different specifications to each other using `+` signs.

The variables that you want to display on the graph must always be wrapped in an `aes()` function, which stands for aesthetics. 
This tells R to determine the value of the aesthetic (x and y axes, colors, groups, line types, etc.) based on the value of the variable. 
`aes()` can be specified both in the main `ggplot()` function (in which case it will apply to all geoms) or within a `geom_...()` function (then it only applies to that geom).

The following code creates a time-series plot of the interest rate in the US over time.
Comments at the end of each line explain what is happening.

```{r}
data_US %>% # feed the previously created data frame into the ggplot function call
  ggplot() + # create the ggplot "blank canvas"
  geom_line(aes(x = year, y = rate)) + # set the variables to be displayed on the x and y axes
  labs(title = "Interest rate in the US", # add a title to the plot
       x = "Year", # set the x axis label
       y = "Interest rate (%)") + # set the y axis label
  theme_light() # change the color scheme and layout of the plot to a different theme
```

If you are happy with how a figure looks, you can save it using the `ggsave()` function and specifying the file name you want to use:

```{r, eval = FALSE}
ggsave("interest_rate.jpg") # save the plot
```

If instead of a time series plot, you want to make a scatterplot, you would use `geom_point()` instead of `geom_line()`.
Additionally, you can use `geom_text()` to display labels next to the points.
This requires you to set an additional aesthetic: the label you want displayed.
Other settings are also possible within each geom, such as size, color, transparency, etc.

```{r}
data_US %>% 
  ggplot(aes(x = GDP, y = rate)) + # if the aesthetics are set in the main ggplot() function, they apply to all geoms
  geom_point() + # make a scatterplot
  geom_text(aes(label = year), nudge_y = 0.25, size = 3) + # add the years as labels next to each point
  labs(title = "Output and interest rate in the US (2006-2012)",
       x = "GDP growth (%)",
       y = "Interest rate (%)") +
  theme_light()
```
```{r, eval = FALSE}
ggsave("scatterplot.jpg") 
```

You can also connect `ggplot` figures to data cleaning/manipulation processes using the `%>%` operator introduced previously.
The code below connects the data manipulation shown previously to a time-series plot of the components of GDP:

```{r}
data_US %>% 
  select(year, consumption, investment, govt_spending) %>% 
  pivot_longer(cols = -year, names_to = "component", values_to = "value") %>%
  ggplot() +
  geom_line(aes(x = year, y = value, color = component)) +
  labs(title = "Components of GDP in the US",
       x = "",
       y = "Percentage change") +
  theme_light()
```
```{r, eval = FALSE}
ggsave("GDP.jpg")
```

# Useful links

You can find additional materials, including useful data sources, and online materials for using R and `ggplot` on the [Data Center website](https://ucrdatacenter.github.io/SSCECON207#Suppporting_materials).
